---
title: "Visualization"
subtitle: Part I
output: github_document
---

```{r setup}
library(tidyverse) #the functions exist in tidyverse but not basic R)
library(ggridges)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"), #3 monitor stations to pull (allow to access data online)
    var = c("PRCP", "TMIN", "TMAX"), # several variables
    date_min = "2021-01-01", #begin date
    date_max = "2022-12-31") |> #end date
  mutate( #reorganize the dataset
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY",
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10, #data manipulation
    tmax = tmax / 10) |> #change the temperature unit to Celcius
  select(name, id, everything()) #select process

weather_df
```

## Scatterplot

Create my first scatterplot ever.
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax )) + #variable map to x feature
  geom_point() #gives the scatterplot
```

Same way to do this

```{r}
weather_df %>% #gives the same plot as above
  ggplot(aes(x =tmin, y = tmax)) +
  geom_point()
```

Another way to interact with `ggplot` (save and edit a plot object)

```{r}
weather_plot =
  weather_df %>% 
  ggplot(aes(x =tmin, y = tmax))

weather_plot + geom_point()
```

## Asvanced scatterplot

**Start with the same one and make it fancy!**
```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax, color = name)) +
  geom_point()
# get better visual information
```

Add a new geometry

```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) #se stands for "standard error"
```
 
What about the `aes` placement?

**What does `geom_smooth` do?**

```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax)) +
  geom_point(aes(color = name)) + #color only apply to the scatterplots but not the smooth line
  geom_smooth()
```

Lect's facet some things!

```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax, alpha = tmin, color = name)) +
  geom_point() + #change the transparency and size of geom_points
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) #grid(rows ~ columns)
#Got single plot panel and smooth curve for each data
```

Let's combine some elements and try a new plot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

## Some small notes

How many geoms have to exist?

> You can have whatever geoms you want.

The `geom_point` is not necessary, the `geom_smooth` can exist along but only gives the smooth curve not the points.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se= FALSE)
```

Import the new package neat geom (`geom_hex`)
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() #Helps understanding the distribution of plots
```

`geom_bin2d` same as above but squared

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_bin2d()
```

`geom_density2d` gives (countor plot) information about elevation
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density2d()
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density2d() +
  geom_point(alpha = 0.3) #Add points
```

**Done with small notes on `ggplot`**

## Unvariate plots

**Histograms** are really great

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Can we add color ..

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

Lest's try a new geometry!

`geom_density` is to smooth down the edge of histogram

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill =name)) +
  geom_density(alpha = 0.3, adjust = 0.5)
```

What about `boxplots`?

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

## Trendy plots

`geom violin`
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = 0.5) +
  stat_summary(fun = "mean")
```

Ridge plots -- the most popular plot of 2017

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```






