---
title: "Visualization"
subtitile: "Part II"
output: github_document
date: "2023-10-10"
---

```{r}
library(tidyverse)
library(patchwork) #Arrange panels
```

## Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"), #3 monitor stations to pull (allow to access data online)
    var = c("PRCP", "TMIN", "TMAX"), # several variables
    date_min = "2021-01-01", #begin date
    date_max = "2022-12-31") |> #end date
  mutate( #reorganize the dataset
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY",
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10, #data manipulation
    tmax = tmax / 10) |> #change the temperature unit to Celcius
  select(name, id, everything()) #select process

weather_df

```

## Remember this plot

**More fine-grained detail on plotting**

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5)
```

## Labels

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  )
```

## Scales

Start with the same plots , a and y scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  scale_x_continuous( #Define scales
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    trans = "sqrt" #Transformation (sqrt, log)
  )
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  scale_x_continuous( #Define scales
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    position = "right"
  )
```

Look at color scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  scale_color_hue ( #Describe features of colors)
    name = "Location",
    h = c(100, 300)) 
```

Use `scale_color_hue` to define your own color
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis ( #Describe features of colors)
    name = "Location",
    discrete = TRUE)
```


## Themes

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis ( #Describe features of colors)
    name = "Location",
    discrete = TRUE) +
  theme(legend.position = "bottom")
```

## Change the whole theme

Popular one `theme_bw`

You get a box around the outside, white background, and grey grid lines

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis ( #Describe features of colors)
    name = "Location",
    discrete = TRUE) +
  theme_bw()
```

`theme_minimal` almost the same with `theme_bw`, but takes the box outside of the graphs.
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis ( #Describe features of colors)
    name = "Location",
    discrete = TRUE) +
  theme_minimal()
```

**Other things to use**

`theme_classic` (no grid lines in the background, *not prefer*)

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis ( #Describe features of colors)
    name = "Location",
    discrete = TRUE) +
  theme_classic()
```

`ggthemes::theme_economist`

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis ( #Describe features of colors)
    name = "Location",
    discrete = TRUE) +
  ggthemes::theme_economist()
```

`ggthemes_theme_excel` (to make it looks like excel)

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis ( #Describe features of colors)
    name = "Location",
    discrete = TRUE) +
  ggthemes::theme_excel()
```

Back to `theme_minimal`

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from rnoaa package; temperatures in 2017."
  ) +
  viridis::scale_color_viridis ( #Describe features of colors)
    name = "Location",
    discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom")
```


##Setting options

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom")) #Update the theme everywhere in this project

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

sacle_color_discrete = scale_color_viridis_d
sacle_fill_discrete = scale_fill_viridis_d
```

## Data args in `geom`

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

molokai =
  weather_df %>% 
  filter(name == "Molokai_HI")

ggplot(data = molokai, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)
```


## `patchwork`

Remember faceting?

```{r}
weather_df %>% 
  ggplot(aes( x= tmin, fill = name)) +
  geom_density(alpha = 0.5) +
  facet_grid(. ~ name)
```

What happens when you want multipanel plots but can't facet?

```{r tmax_tmin_plot}
tmax_tmin_plot =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  theme(legend.position = "none")

prcp_dens_plot =
  weather_df %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5)
```

Add filtering to `prcp_dens_plot`

```{r filtered prcp_dens_plot}
prcp_dens_plot =
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) +
  theme(legend.position = "none")
```

Seasonality

```{r tmax_date_plot}
tmax_date_plot =
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none")
```

Put the 3 plots together

```{r}
# Give the first plot more space
tmax_tmin_plot + (prcp_dens_plot + tmax_date_plot)

tmax_tmin_plot / (prcp_dens_plot + tmax_date_plot)

(tmax_tmin_plot + prcp_dens_plot) / tmax_date_plot
```

## Data manipulation

Control your factors.

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

To add *factors* to the plot

```{r}
weather_df %>% 
  mutate(
    name = factor(name)
  ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

To change orders of the three plots by adding *factors*

```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = forcats::fct_relevel(name, c("Molokai_HI"))
  ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

What if I wanted densities for tmin and tmax simultaneously? (want tmin and tmax overlad on top of each other)

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% #data tidying
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperatures"
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = 0.5)
```

More about data tidiness issue

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperatures"
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = 0.5) +
  facet_grid(. ~ name)
```

Quick example

```{r}

```







